Demographic estimates from age data

### 2021-01-19 manuscript.Rmd
parsimonious model of age distribution

More community based delivery (Ahmed's paper, Pivot, accessibility, etc.) Possibly combining different campaign strategies could work per models Turnover is tricky & waning doesn't contribute too much. Thermotolerant & 3 yrs even without booster. (although most dogs are revaccinated and we do this in the model) Volunteer based/NGO based. Push for govt investment and viewing as a public good. The veterinary workforce (relieving pressures on them, challenges with private-public roles).

>>>>

More community based delivery (Ahmed's paper, Pivot, accessibility, etc.) Possibly combining different campaign strategies could work per models Turnover is tricky & waning doesn't contribute too much. Thermotolerant & 3 yrs even without booster. (although most dogs are revaccinated and we do this in the model). Volunteer based/NGO based. Push for govt investment and viewing as a public good. Modeling after childhood vax = routine through local offices of puppies + supplemental activities by trained vets annually to catch unvax (feral, oral). The veterinary workforce (relieving pressures on them, challenges with private-public roles).

### 2021-01-19 qs_for_jess/demography.Rmd
parsimonious model of age distribution

# pkgs
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(patchwork)
library(cowplot)

@@

```{r}
# Age pyramid
animals_age <- read_csv("animals_age.csv")

@@

```{r}
agemod1 <- function(par = c(psurv_pup = -1, 
                            psurv_adult = -1),
                    fert = 3, 
                    n_per_age, type = "fit") {

@@

@@

    # balance births (all to susceptibles)
    S_pup[1, t] <- births # get replaced

@@

    # do puppy vaccination 
    V_pup[pup_month_vacc, t] <- pups_to_vacc * pup_cov 
    S_pup[pup_month_vacc, t] <- S_pup[pup_month_vacc, t] - pups_to_vacc * pup_cov

@@

    # Vaccinate as part of campaign accounting for revax
    cov_now <- ifelse(t %in% campaign_times, campaign_cov, 0)
    camp_vacc_adult <- (S_adult[t] + revacc * sum(V_adult[, t])) * cov_now - revacc * sum(V_adult[, t])
    camp_vacc_adult <- ifelse(camp_vacc_adult < 0, 0, camp_vacc_adult)

    # every vax is revaccinated with a certain revacc prob
    if(camp_vacc_adult > 0) {
      V_adult[1, t] <- camp_vacc_adult + revacc * sum(V_adult[, t])
      V_adult[2:36, t] <- (1 - revacc) * V_adult[2:36, t]
    } else {
      V_adult[1, t] <- camp_vacc_adult # newly vaccinated
    }

    # Balance addtl campaign vax adults
    S_adult[t] <- S_adult[t] - camp_vacc_adult

    # Balance addtl campaign vax pups (assuming that no puppies are revax)
    camp_vacc_pup <- S_pup[pup_month_vacc:12, t] * cov_now 
    V_pup[pup_month_vacc:12, t] <- V_pup[pup_month_vacc:12, t] + camp_vacc_pup
    S_pup[pup_month_vacc:12, t] <- S_pup[pup_month_vacc:12, t] - camp_vacc_pup

    # Add in surviving pups aging to adults so that they can wane appropriately
    t_to_sub <- pup_month_vacc - 1
    V_adult[13 - t_to_sub, t] <- V_adult[13, t] +  pup_surv * V_pup[12, t - 1]

@@

# for population of 1000
# with campaigns only
camp <- simvacc(tmax = 12*5, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
        revacc = 1, 
        pup_cov = 0, 
        campaign_cov = 0.7, campaign_times = seq(12, 12*5, by = 12), 
        pup_month_vacc = 1, 
        birthrate = 3/12, # monthly birth rate
        S_adult_0 = prop_adults*1000, 
        S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
        V_pup_0 = rep(0, 12))

# with pup vacc only
pup <- simvacc(tmax = 12*5, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
        revacc = 1, 
        pup_cov = 0.7, 
        campaign_cov = 0, campaign_times = 0, 
        pup_month_vacc = 3, 
        birthrate = 3/12, # monthly birth rate
        S_adult_0 = prop_adults*1000, 
        S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
        V_pup_0 = rep(0, 12))

both <- simvacc(tmax = 12*5, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
        revacc = 1, 
        pup_cov = 0.7, 
        campaign_cov = 0.7, campaign_times = seq(12, 12*5, by = 12), 
        pup_month_vacc = 1, 
        birthrate = 3/12, # monthly birth rate
        S_adult_0 = prop_adults*1000, 
        S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
        V_pup_0 = rep(0, 12))

vacc_comp <- bind_rows(tibble(month = 1:60, 
                              cov = camp$V/(camp$S + camp$V), type = "Campaign"), 
                       tibble(month = 1:60, 
                              cov = pup$V/(pup$S + pup$V), type = "Puppy vax"), 
                       tibble(month = 1:60,
                              cov = both$V/(both$S + both$V), type = "Both"))

>>>>

# pkgs
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(patchwork)
library(cowplot)
library(here)

@@

```{r}
# Age pyramid
animals_age <- read_csv(here("qs_for_jess/animals_age.csv"))

@@

```{r}
# constraining growth
agemod1 <- function(par = c(psurv_pup = -1, 
                            psurv_adult = -1),
                    fert = 3, 
                    n_per_age, type = "fit") {

@@

  # Order of steps:
  # 1. all adults have babies (S + V)
  # 2. survival probabilities are applies to everyone
  # 3. Surviving vaccinated indiviudals immunity wanes
  # 4. Then new vax and revax happends

@@

    # balance births (all surviving to susceptibles)
    S_pup[1, t] <- births

@@

    # do puppy vaccination 
    # CHECK: this should be additive no? Or is it always zero? Yes because no vaccs at t yet! Unless others have been vaccinated in campaign?
    V_pup[pup_month_vacc, t] <- pups_to_vacc * pup_cov 
    S_pup[pup_month_vacc, t] <- S_pup[pup_month_vacc, t] - pups_to_vacc * pup_cov

@@

    # Campaign vaccination of adults accounting for revaccination --------------

    # Do the waning (adults only)
    V_adult[2:36, t] <- adult_surv * V_adult[1:35, t - 1]
    # After applying survival probabilities
    S_adult[t] <- S_adult[t] + V_adult[36, t - 1]

    # Coverage now
    cov_now <- ifelse(t %in% campaign_times, campaign_cov, 0)

    # First get the number of currently vaccinated doggoes that are revaccinated
    # Assume that these are distributed evenly across the waning continuum 
    # i.e. same prop across V_adult[, t]
    total_vax_adult <- cov_now * (S_adult[t] + sum(V_adult[, t]))
    revax_adult <- sum(revacc * V_adult[, t])

    # constrain new vaccination by preferentially revaccinating already vaxed adults
    new_vax_adult <- ifelse(revax_adult > total_vax_adult, 0, total_vax_adult - revax_adult)

    # constrain revaccination so it cannot exceed coverage level
    # if greater than total_vax_adult, then adjusted coverage should be total_vax_adult/revac_adult
    # to adjust to total expected vaccinations
    adj_cov <- ifelse(revax_adult > total_vax_adult, 
                      total_vax_adult/revax_adult * V_adult[, t], cov_now * V_adult[, t])

    # subtract these from currently vaccinated adults 
    V_adult[2:36, t] <- V_adult[2:36, t] - adj_cov

    # and move them into the first waning step along with the newly vaccinated
    V_adult[1, t] <- sum(adj_cov) + new_vax_adult

    # Balance the susceptibles as well (subtract newly vaccinated)
    S_adult[t] <- S_adult[t] - new_vax_adult

    # Do the same vax steps for pups
    # But assuming that puppies aren't revaccinated
    camp_vacc_pup <- S_pup[pup_month_vacc:12, t] * cov_now 
    V_pup[pup_month_vacc:12, t] <- V_pup[pup_month_vacc:12, t] + camp_vacc_pup
    S_pup[pup_month_vacc:12, t] <- S_pup[pup_month_vacc:12, t] - camp_vacc_pup

    # Add in surviving pups aging to adults so that they can wane appropriately
    # for example if vaccinated at three months, will enter adulthood at waning stage of 10 months
    # i.e. having spent 9 mos. vaccinated!
    v_to_sub <- 13 - pup_month_vacc
    V_adult[v_to_sub, t] <- V_adult[v_to_sub, t] +  pup_surv * V_pup[12, t - 1]

@@

# for population of 1000
# with campaigns only
camp <- simvacc(tmax = 12*10, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
                revacc = 1, 
                pup_cov = 0, 
                campaign_cov = 0.7, campaign_times = seq(12, 12*10, by = 12), 
                pup_month_vacc = 1, 
                birthrate = 3/12, # monthly birth rate
                S_adult_0 = prop_adults*1000, 
                S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
                V_pup_0 = rep(0, 12))

# with pup vacc only
pup <- simvacc(tmax = 12*10, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
               revacc = 1, 
               pup_cov = 0.7, 
               campaign_cov = 0, campaign_times = 0, 
               pup_month_vacc = 3, 
               birthrate = 3/12, # monthly birth rate
               S_adult_0 = prop_adults*1000, 
               S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
               V_pup_0 = rep(0, 12))

both <- simvacc(tmax = 12*10, pup_surv = surv$psurv_pup, adult_surv = surv$psurv_adult, 
                revacc = 1, 
                pup_cov = 0.7, 
                campaign_cov = 0.7, 
                campaign_times = seq(12, 12*10, by = 12), 
                pup_month_vacc = 1, 
                birthrate = 3/12, # monthly birth rate
                S_adult_0 = prop_adults*1000, 
                S_pup_0 = rep(prop[1]/12, 12)*1000, V_adult_0 = 0, 
                V_pup_0 = rep(0, 12))

vacc_comp <- bind_rows(tibble(month = 1:120, 
                              cov = camp$V/(camp$S + camp$V), type = "Campaign"), 
                       tibble(month = 1:120, 
                              cov = pup$V/(pup$S + pup$V), type = "Puppy vax"), 
                       tibble(month = 1:120,
                              cov = both$V/(both$S + both$V), type = "Both"))
